Verifying Spatial Properties of Array Computations

### 2017-04-12 oopsla17/oopsla17.tex
moving the model into the main text and adapting

@@

We used Wheeler's \emph{SLOCCount} to get a count of the physical
source lines (excluding comments and blank
lines)~\cite{wheeler2001sloccount}; including them our corpus
totals

@@

\begin{defn}[Extended integers]
  We define the set \zinf{} as
  $\mathbb{Z}$ extended with $\infty$ and $-\infty$. For any $a$ in
  \zinf{}, we have $-\infty \leq a \leq \infty$. The resulting set is
  a total order with top and bottom elements $-\infty$ and
  $\infty$ respectively.
\end{defn}

\begin{defn}[Index scheme]
  An $N$-dimensional \emph{index scheme} is a vector of length $N$ of
  extended integer sets, \ie{}, a member of $\mathcal{P}(\zinf)^{N}$.
  An alternate (equivalent) view is of a $N$-times finite Cartesian
  product on subsets of $\zinf$.  We use $S, T, U$ to denote index
  schemes.

Index schemes can be \emph{projected} in the $i^{\textit{th}}$ dimension by
  $\pi_i : {\mathcal{P}(\zinf{})}^N \to \mathcal{P}(\zinf{})$ where $1 \leq i
  \leq N$. For an index scheme $I$, we refer to $\pi_i(S)$ as the
  \emph{$i^{th}$ component} of $S$.
\end{defn}

@@

\begin{lem}{}\label{lem:vector-intersect}
  Intersection distributes over $N$-dimensional index schemes. 
  That is, for index schemes $I, J \in \mathcal{P}(\zinf)^N$ then:
%
  \begin{equation*}
    I \cap J = \prod_{i = 1}^{N} \pi_i(I) \cap \pi_i(J)
  \end{equation*}
\end{lem}

@@

\begin{lem}{}\label{lem:vector-union}
Let $I, J$ be $N$-dimensional index schemes such that
$\pi_i(I) = \pi_i(J)$ for all $1 \leq i \leq N$ apart 
from some dimension $k$ where $\pi_k(I) \neq \pi_k(J)$,
then:
%
  \begin{equation*}
    I \; \cup \; J
    =
    \pi_1(I) \times \cdots \times
    (\pi_k(I) \cup \pi_k(J)) \times \cdots \times
    \pi_N(I)
  \end{equation*}
%
\end{lem}
%
\paragraph{Intervals}

We define an extended notion of closed interval on

\begin{defn}{(\emph{Interval subspace})}
  An interval subspace is a finite Cartesian product of intervals on \zinf{},
  denoted by the set $\textit{Interval}^N$ for a product of $N$ intervals. An
  interval subspace is an offset subspace.
\end{defn}

@@

\begin{defn}
The function $\textit{schematic}_I$ parameterised by
a set of induction variables $I$, which maps
from an array subscript into a $n$-dimensional (closed) interval:
%
\begin{align*}
\textit{schematic}_I(\bar{e}) & =
\prod_{i \in |\bar{e}|} \textit{interval}(e_i)
\qquad
\textit{interval}_I(e) = \begin{cases}
[c, c] & e = i \pm c \, \wedge \, i \in I \\
[-\infty, \infty] & FV(e) \cap I = \emptyset
\end{cases}
\end{align*}
Note that \textit{inveral} may be undefined (it is a partial function)
in which case \textit{schematic} is undefined if any of its
components are undefined.
\end{defn}

>>>>

\usepackage{thmtools}
\usepackage{thm-restate}

@@

We used Wheeler's \emph{SLOCCount} to get a count of the physical
source lines (excluding comments and blank
lines)~\cite{wheeler2001sloccount}; including non-physical lines
 our corpus totals \dnote{insert}.

@@

%\begin{defn}[Extended integers]
%  We define the set \zinf{} as
%  $\mathbb{Z}$ extended with $\infty$ and $-\infty$. For any $a$ in
%  \zinf{}, we have $-\infty \leq a \leq \infty$. The resulting set is
%  a total order with top and bottom elements $-\infty$ and
%  $\infty$ respectively.
%\end{defn}

\begin{defn}[Index scheme]
  An $N$-dimensional \emph{index scheme} is a vector of length $N$ of
  extended integer sets, \ie{}, a member of $\mathcal{P}(\bz)^{N}$.
  An alternate (equivalent) view is of a $N$-times finite Cartesian
  product on subsets of $\bz$.  We use $S, T, U$ to denote index
  schemes.

Index schemes can be \emph{projected} in the $i^{\textit{th}}$ dimension by
  $\pi_i : {\mathcal{P}(\bz{})}^N \to \mathcal{P}(\bz{})$ where $1 \leq i
  \leq N$. For an index scheme $I$, we refer to $\pi_i(S)$ as the
  \emph{$i^{th}$ component} of $S$.
\end{defn}

@@

\begin{restatable}{lem}{vectorIntersect}
\label{lem:vector-intersect}
  Intersection distributes over $N$-dimensional index schemes. 
  That is, for index schemes $S, T \in \mathcal{P}(\bz)^N$
%
  \begin{equation*}
    S \cap T = \prod_{i = 1}^{N} \pi_i(S) \cap \pi_i(T)
  \end{equation*}
\end{restatable}

@@

\begin{restatable}{lem}{vectorUnion}
\label{lem:vector-union}
Let $S, T$ be $N$-dimensional index schemes such that
$\pi_i(S) = \pi_i(T)$ for all $1 \leq i \leq N$ apart 
from some dimension $k$ where $\pi_k(S) \neq \pi_k(T)$,
then:
%
  \begin{equation*}
    S \; \cup \; T
    =
    \pi_1(S) \times \cdots \times
    (\pi_k(S) \cup \pi_k(T)) \times \cdots \times
    \pi_N(S)
  \end{equation*}
%
\end{restatable}
%

\begin{defn}
  We define an extended notion of closed interval on $\bz{}$ which may contain
  a \emph{hole} at the origin, written \interv{a}{b}{c} where $a$ and $b$ are
  drawn from \bz{} with $a \leq 0 \leq b$
   and $c$ is drawn from $\mathbb{B} = \{ \mathsf{true}, \mathsf{false}
  \}$. Intervals are interpreted as sets as follows:
%
  \begin{equation*}
    \interv{a}{b}{c} \triangleq
      \{ n \mid a \leq n \leq b \wedge (\neg c \implies n \neq 0) \}
  \end{equation*}
%
  We denote the set of all such intervals (sets) as
   $\textit{Interval}$. If the superscript is
  omitted it is treated as $\mathsf{true}$.
\end{defn}

\begin{restatable}{lem}{intervalIdentities}
 \label{lem:interval-identities}
  We have the following dual identities for \bz{} intervals:
%
  \begin{align*}
    \interv{a}{b}{c} \cap \interv{d}{e}{f} & =
      \interv{\max \{a,d\}}{\min \{b,e\}}{c \wedge f} \\
    \interv{a}{b}{c} \cup \interv{d}{e}{f} & =
      \interv{\min \{a,d\}}{\max \{b,e\}}{c \vee f}
  \end{align*}
\end{restatable}
%

\begin{defn}{(\emph{Interval subspace})}
  An interval subspace is a finite Cartesian product of intervals on
  \bz{} or \bz itself, 
  denoted by the set $\textit{Interval}^N$ for a product of $N$ intervals.
\end{defn}

@@

\begin{defn}
The function $\textit{schematic}_I$ parameterised by
a set of induction variables $I$, which maps
from an array subscript into a $n$-dimensional (closed) interval:
%
\begin{align*}
\textit{schematic}_I(\bar{e}) & =
\prod_{i \in |\bar{e}|} \textit{interval}(e_i)
\qquad
\textit{interval}_I(e) = \begin{cases}
[c, c] & e = i \pm c \, \wedge \, i \in I \\
\bz & FV(e) \cap I = \emptyset
\end{cases}
\end{align*}
Note that \textit{inveral} may be undefined (it is a partial function)
in which case \textit{schematic} is undefined if any of its
components are undefined.
\end{defn}
